---
title: "Data Preparation and Principle Component Analysis"
author: "OEB"
date: "March 1, 2018"
output: 
  html_document: 
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE )
```

```{r libs}

suppressPackageStartupMessages( require(oetteR) ) 
suppressPackageStartupMessages( require(tidyverse) )

```


# Cleaning Data

Before I learned about `recipes`. I had my own set of functions that would do something similar with less functionality of course. `oetteR::f_clean_data()` takes a dataframe and performs some automated cleaning steps and sorts the variables into categrocial and numerical categories and returns a list which I always name `data_ls`.


```{r clean}

data = ISLR::Auto

data_ls = f_clean_data( data = data
                         # reduce number of levels to 10, group to other
                         , max_number_of_levels_factors = 10
                         # numericals with less than 10 unique values will be converted to factors
                         , min_number_of_levels_nums = 10
                         # exclude missing data
                         , exclude_missing = T
                         # negative values will be set to zero
                         ,replace_neg_values_with_zero = T
                         # allow negative values in these columns
                         ,allow_neg_values = 'null'
                         # tag id columns
                         , id_cols = 'name'
                         )


print( str(data_ls) )


```

Notice that numericals are converted to ordered factors

# BoxCox Transformation

We can use a boxcox transformation on numerical variables.


```{r}

data_ls = f_boxcox(data_ls)

print( str(data_ls) )

```

# PCA

Base R has a decent function for PCA `prcomp()`. However the returned object is a bit messy and contains a lot of matrices intesad of dataframes. `oetteR::f_pca()` is a convenience wrapper which uses `data_ls` lists. 


```{r}

pca_ls = f_pca( data_ls
             , center = T
             , scale = T
             , use_boxcox_tansformed_vars = T
             , include_ordered_categoricals = T
             )


```

The returned pca object has some new features compared to `prcomp()`

## Call the original Data with the Principle Components as additional Variables

```{r}

as_tibble( pca_ls$data )

```

## Get variance explained as a dataframe

```{r}

pca_ls$pca$vae

```

## How much is each variable contributing to each component in percent

Note the columns add up to 100

```{r}

pca_ls$pca$contrib

```

## Contribution of variables to variance explained by principle component

```{r}

pca_ls$pca$contrib_abs_perc


```

# Plot PCA

## Plot variance explained

Note returns a plotly graph by default

`oetteR::f_pca_plot_variance_explained`

```{r}

f_pca_plot_variance_explained(pca_ls
                              # dont include componetns that explain less than 2.5 percent of the variabnce
                              , threshold_vae_for_pc_perc = 2.5)


```

## Plot first two principle components and color by cylinder

and color cylinders, `oetteR::f_pca_plot_variance_explained` returns a taglist created by `htmltools::tagList` which stores a `plotly` graph and a `DT:datatable`. The algebraic sign (+/-) of the rotation value in the table tells you whether a contribution of one variable to one principle component is positive or negative.



```{r}

taglist = f_pca_plot_components( pca_ls
                                 , x_axis = 'PC1'
                                 , y_axis = 'PC2'
                                 , group = 'cylinders' )

taglist

```

