---
title: "GGpubr"
author: "OEB"
date: "January 6, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`ggpubr` is a package that offers a sophisticated interface to produce nice histograms and violin plots for numeric variables and easily ads significance values to the plots. `oetteR::f_plot_hist` takes a data_ls list and plots a histogram or a bar plot of frequencies of all variables but the histograms are lacking some of `ggpubr` features, which are not so easily replicated. Here we want to try whether we can modify `oetteR::fplot_hist` to use most of `ggpubr` features


```{r}
suppressPackageStartupMessages( require(oetteR) )
suppressPackageStartupMessages( require(ggpubr) )
suppressPackageStartupMessages( require(tidyverse) )

```

# Data

```{r}

data_ls = ISLR::Auto %>%
  f_clean_data()

```

# Density Histograms

```{r}

ggdensity( data = data_ls$data
           , x = 'displacement'
           , y = '..count..'
           , color = 'cylinders'
           , fill = 'cylinders'
           , palette = f_plot_col_vector74()
           , rug = T
           , add = 'mean'
           )


f_plot_hist( data_ls
             , variable = 'displacement'
             , group = 'cylinders'
             , y_axis = 'count'
             , graph_type = 'line'
             , rug = T )
```

# Bar Histograms

```{r}

gghistogram( data = data_ls$data
           , x = 'displacement'
           , y = '..count..'
           , color = 'cylinders'
           , fill = 'cylinders'
           , palette = f_plot_col_vector74()
           , rug = T
           , add = 'mean'
           )


f_plot_hist( data_ls
             , variable = 'displacement'
             , group = 'cylinders'
             , y_axis = 'count'
             , graph_type = 'bar'
             , rug = T )
```

# Violin Plots

```{r}

v1 = ggviolin( data = data_ls$data
           , x = 'cylinders'
           , y = 'displacement'
           #, color = 'cylinders'
           , fill = 'cylinders'
           , palette = f_plot_col_vector74()
           , rug = T
           , add = 'boxplot'
           , add.params = list( fill = 'white')
           )

v1

v2 = f_plot_hist( data_ls
             , variable = 'displacement'
             , group = 'cylinders'
             , y_axis = 'count'
             , graph_type = 'violin'
             , rug = T )

v2

```


# Add stats

## all combinations

```{r}

lvl = levels(data_ls$data$cylinders) 

compare = expand.grid( a = lvl
                       , b = lvl
                       , stringsAsFactors = F) %>%
  filter( a != b) %>%
  mutate( comb = map2( a, b, function(x,y) sort( c(x,y) ) )
          , comb = map( comb, paste0, collapse = ',' )
          ) %>%
  unnest( comb ) %>%
  group_by( comb ) %>%
  summarise() %>%
  mutate( comb = stringr::str_split(comb, ',') ) %>%
  .$comb

```


```{r error=FALSE }

v1_1 = v1 + 
  stat_compare_means(comparisons = compare, label = "p.signif") 

v1_1


ggplot( data_ls$data, aes( cylinders
                           , displacement
                           , fill = cylinders) ) +
  geom_violin() + 
  stat_compare_means(comparisons = compare, label = "p.signif") 


p = f_plot_pretty_points( df = data_ls$data
                          , col_y = 'displacement'
                          , col_x = 'weight'
                          , col_facet = 'origin') +
  geom_smooth() + 
  ggpubr::stat_cor()

p

```

# Hide NS values

Make random grouping

```{r}

data_ls$data$rnd = as.factor( sample( LETTERS[1:5], nrow(data_ls$data), replace = T ) )

```

filter all pairs that have a t_test p_val over 0.05

```{r}

f_compare_means = function(comb, data, col_var, col_group){
  
  sym_group = as.name(col_group)
  
  lvl1 = stringr::str_split(comb, ',')[[1]][1]
  lvl2 = stringr::str_split(comb, ',')[[1]][2]
  
  x = data %>%
    filter( rlang::UQ(sym_group) == lvl1 ) %>%
    .[[col_var]]
  
  y = data %>%
    filter( rlang::UQ(sym_group) == lvl2 ) %>%
    .[[col_var]]
  
  t_test = t.test(x,y)
  
  return( t_test$p.value )
}

lvl = levels(data_ls$data$rnd) 

compare_filt = expand.grid( a = lvl
                       , b = lvl
                       , stringsAsFactors = F) %>%
  filter( a != b) %>%
  mutate( comb = map2( a, b, function(x,y) sort( c(x,y) ) )
          , comb = map( comb, paste0, collapse = ',' )
          ) %>%
  unnest( comb ) %>%
  group_by( comb ) %>%
  summarise() %>%
  mutate( p_val  = map_dbl( comb, f_compare_means
                            , data = data_ls$data
                            , col_var = 'displacement'
                            , col_group = 'rnd' )
          , comb = stringr::str_split(comb, ',') ) %>%
  filter( p_val <= 0.05 ) %>%
  .$comb

```


```{r}

ggviolin( data = data_ls$data
           , x = 'rnd'
           , y = 'displacement'
           , color = 'rnd'
           , fill = 'rnd'
           , palette = f_plot_col_vector74()
           , rug = T
           , add = 'boxplot'
           , add.params = list( fill = 'white')
           ) + 
  stat_compare_means(comparisons = compare, label = "p.signif") 


ggviolin( data = data_ls$data
           , x = 'rnd'
           , y = 'displacement'
           , color = 'rnd'
           , fill = 'rnd'
           , palette = f_plot_col_vector74()
           , rug = T
           , add = 'boxplot'
           , add.params = list( fill = 'white')
           ) + 
  stat_compare_means(comparisons = compare_filt, label = "p.signif") 


```


# plotly compatibility

```{r}

v1_1 %>%
  plotly::ggplotly()

v1 %>%
  plotly::ggplotly()

v2 %>%
  plotly::ggplotly()

ggdensity( data = data_ls$data
           , x = 'displacement'
           , y = '..count..'
           , color = 'cylinders'
           , fill = 'cylinders'
           , palette = f_plot_col_vector74()
           , rug = T
           , add = 'mean'
           ) %>%
  plotly::ggplotly()

p %>%
  plotly::ggplotly()

```


# Summary

`ggpubr` is a great to add p values to any ggplot object and the histograms do look better than the ones from my default function. Unfortunately the p values dont show in plotly. `ggpubr` also does not plot frequency distributions of factor variables and does not offer a stat_ function to add a chi square p value. 







