---
title: "mix taglist and plots"
author: "OEB"
date: "February 6, 2018"
output: 
  html_document: 
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

```{r}

suppressPackageStartupMessages( require(oetteR) )
suppressPackageStartupMessages( require(tidyverse) )

```


Presently I am only printing lists of one type of printable object as html using `f_plot_obj_2_html`. However I would like to mix these lists. Having plotly graphs right next to regular ggplot graphs. 

# Generate Mixed list

```{r}

tab = f_datatable_universal( mtcars )

p = ggplot(mtcars, aes(disp, mpg) ) +
  geom_point()

tag1 = f_html_padding( tab, title = 'tag1')

tag2 = f_html_padding( tab, title = 'tag2') 

p_plotly = plotly::ggplotly(p)

tabplot = tabplot::tableplot( mtcars, plot = F )

mix = list( p_plotly, tag1, p, tabplot, tab) 


```


# Try to identify object

```{r results='asis'}

objects = tibble( obj = mix ) %>%
  mutate( class = map(obj, class)
          , class = map_chr( class, function(x) x[[1]] )
          , print_f = map(class, function(x) ifelse( x == 'tabplot',tabplot:::plot.tabplot, print) )
          , list_type = map_chr(class, function(x) ifelse( x %in% c('plotly', 'shiny.tag.list', 'datatables')
                                                     , 'taglist', 'regular_list') )
          )

objects_marked = objects %>%
  mutate(  rank = rank(list_type)
          , change_lag = lag(rank)
          , change_event = rank != change_lag
          , change_num = ifelse( is.na(change_event) | change_event == FALSE, 0, 1 )
          , cum_change = cumsum(change_num) 
          )


objects_collapse = objects_marked %>%
  group_by( list_type, cum_change ) %>%
  nest() %>%
  arrange( cum_change )


print_obj = function( tib, list_type){
  
  
  if( list_type == 'taglist' ){
    
    htmltools::tagList( tib$obj )
    
  }else{
    
    walk2( tib$obj, tib$print_f, function(x,y) y(x) ) 
    
  }
  
}


for( p in list(p_plotly, p_plotly) ){
  
  htmlwidgets:::knit_print.htmlwidget(p) # does not work inside loops or functions
  
}


walk2( objects_collapse$data, objects_collapse$list_type, print_obj )


```

it seems impossible to print a plotly object or a DT::datatable or a taglist inside a function or a loop. We could either try to built a custom print function with knitr.

https://cran.r-project.org/web/packages/knitr/vignettes/knit_print.html

which requires understanding of R objects.

or we could merge different Rmd files before rendering which seems a bit too hacky. I think its best to simply seperate htmlwidgets/taglists and other printable objects for the time being.


