---
title: "Prediction Intervalls"
author: "OEB"
date: "December 21, 2017"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
    theme: sandstone
    df_print: kable
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, results = 'hide')
```


```{r}

suppressPackageStartupMessages( require(tidyverse) )
suppressPackageStartupMessages( require(oetteR) )

```

# Various Intervalls of a normally distributed sample

```{r}

map_intervalls_norm = function( .f = rnorm, n = 50, ...){

  set.seed(1)
  
  pop = .f(n, ...)
  
  me = mean(pop)
  std = sd(pop)
  sem = std / sqrt( length(pop) )
  
  df = tibble( value = c('mean','SEM','2xSEM(CI95)','SD','2xSD(PI95)')
               , min = c( me, me - sem, me - 2 * sem, me - std, me - 2 * std )
               , max = c( me, me + sem, me + 2 * sem, me + std, me + 2 * std )
               , pop = list(pop) ) %>%
    mutate( value = fct_relevel(value, 'mean','SEM','2xSEM(CI95)','SD') )
  
  return(df)

}


plot_intervals = function(df, title){
  
  df_unnest = df %>%
    unnest(pop) %>%
    mutate( fill = case_when( pop < min | pop > max ~ 'out'
                              , TRUE ~ 'in' ) )
  
  p = ggplot(df) +
    geom_errorbar( aes(x = value, ymin = min, ymax = max)
                   , size = 1) +
    ggbeeswarm::geom_beeswarm( aes(x = value, y = pop, fill = fill)
                               , data = df_unnest
                               , shape = 21) +
    theme( legend.position = 'none') +
    geom_hline(yintercept = 0, color = 'grey', linetype = 2, size = 1 ) +
    labs( x = '', y = '', subtitle = title )
  
  return(p)

}

title = paste( c('calculated intervalls for 50 samples taken from a normal distributed population'
                 , 'with mean = 0 and sd = 1'), collapse = '\n' )

df = map_intervalls_norm( .f = rnorm)

plot_intervals(df, title = title)



```

# Misleading Intervalls if sample is not normally distributed

```{r}

title = paste( c('calculated intervalls for 50 samples taken from a gamma distributed population'
                 , 'with rate = 1 and scale = 1/rate'), collapse = '\n' )

df = map_intervalls_norm( .f = rgamma, shape = 1)

plot_intervals(df, title = title)


```

# Bootstrap method can be used to calculate intervalls for none-normally distributed samples

```{r}


map_intervalls_boot = function(){

  require(boot)
  
  set.seed(1)
  
  pop = rgamma(50, shape = 1)
  
  boundary = c(0.975, 0.84, 0.5, 0.16, 0.025)
  
  sim_mean = boot( pop
                , function(data, index) mean(data[index])
                , 1000 )
  
  sim_mean = sim_mean$t
  
  me_boot = quantile(sim_mean, boundary)
  
  df_boot = tibble(boundary = boundary )%>%
    mutate( boot = map( boundary, function(x) boot(pop, function(data, index) quantile(data[index], x), 1000 ) )
            , boot_value = map_dbl(boot, 't0') ) 
  
  
  df = tibble( value = c('mean','SEM','2xSEM(CI95)','PI68','PI95')
               , min = c( me_boot[3], me_boot[4], me_boot[5], df_boot$boot_value[4], df_boot$boot_value[5] )
               , max = c( me_boot[3], me_boot[2], me_boot[1], df_boot$boot_value[2], df_boot$boot_value[1] )
               , pop = list(pop) ) %>%
    mutate( value = fct_relevel(value, 'mean','SEM','2xSEM(CI95)','PI68','PI95') )

}

df = map_intervalls_boot()

title = paste( c('calculated intervalls for 50 samples taken from a gamma distributed population'
                 , 'with rate = 1 and scale = 1/rate using bootstrap'), collapse = '\n' )


plot_intervals(df, title = title)

```

# Calculating prediction intervalls for linear regression

When we use a linear model to predict a none-linear context the residuals are not normally distributed throughout the prediction space.

```{r}

set.seed(1)

m = lm(price~carat, ggplot2::diamonds)

df = tibble(pred = predict(m)
            , resid = m$residuals
            , obs = ggplot2::diamonds$price
            , min = obs - predict(m, se.fit = T)$se.fit * 2
            , max = obs + predict(m, se.fit = T)$se.fit * 2 ) %>%
  mutate( fill = case_when( obs < min | obs > max ~ 'out'
                              , TRUE ~ 'in' ) ) %>%
  sample_n(250)

df %>%
  select(pred, obs, resid) %>%
  gather( value = 'value', key = 'key', - pred) %>%
  ggplot( aes(x = pred, y = value) ) +
    geom_point( ) +
    geom_smooth( method = 'lm') +
    facet_wrap(~key)



```

We can seperate the prediction space into bins and use the bootstrap method to calculate the prediction intervalls for each bin

Here we bin 250 value pairs into 10 equally sized bins of 25 observations

```{r}

df_pi = f_prediction_intervall_raw(df, 'pred', 'obs'
                                   , bootstrap = T
                                   , n_neighbours = 25
                                   , intervall = 0.975) %>%
  f_prediction_intervall_raw('pred', 'obs'
                               , bootstrap = T
                               , n_neighbours = 25
                               , intervall = 0.025)


ggplot(df_pi) +
  geom_ribbon( aes(x = pred
                   , ymax = pred_PI97.5_raw
                   , ymin = pred_PI2.5_raw) ) +
  geom_point(aes(x = pred, y = obs, color = steps)
             , show.legend = F) +
  scale_color_manual( values = f_plot_col_vector74(greys = F))

```

Here we bin 53940 obseravations into 107 bins of 500 observations each.

Only a subset of 500 random points will be plotted

```{r}

m = lm(price ~ carat + depth, ggplot2::diamonds)

df = tibble( obs = ggplot2::diamonds$price
           , pred = predict(m, newdata = ggplot2::diamonds) ) %>%
   f_prediction_intervall_raw( 'pred','obs', intervall = 0.975) %>%
   f_prediction_intervall_raw( 'pred','obs', intervall = 0.025)

p = f_plot_pretty_points( dplyr::sample_n(df, 500), 'pred', 'obs' ) +
  geom_ribbon(mapping = aes(ymin = pred_PI2.5_raw
                            , ymax = pred_PI97.5_raw
                            , fill = 'lightsalmon')
              , data = df
              , color = NA
              , fill = 'lightsalmon'
              , alpha = 0.3 ) + 
  geom_line(mapping = aes(x = pred
                          , y = pred_mean_raw
                          )
            , data = df
            , color = 'deepskyblue4'
            #, size = 1
            ) +
  coord_cartesian( xlim=c(0,20000), ylim = c(0,20000))

# reorganize layers
p$layers = list(p$layers[[2]], p$layers[[1]], p$layers[[3]] )

p

```


