---
title: "Content"
author: "Bjoern Oettinghaus"
date: "27 Dezember 2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```


```{r}

suppressPackageStartupMessages( require(tidyverse) )
suppressPackageStartupMessages( require(oetteR) )

```

# Github

[link to the oetteR github page](https://github.com/erblast/oetteR)

here you can find an introduction to the package functions and links to rendered versions of the vignettes and rendered versions of most of the Rmd POCs

# Vignettes

These vignettes give you an overview about what functions are in the package and how you can use them

```{r}

wd = file.path( system.file(package = 'oetteR') )
folder_Rmd = 'vignettes'
folder_html = 'vignettes'

f_html_table_html_and_rmd_link( wd, folder_Rmd, folder_html )

```

# POC Rmd

These are some proof of concepts that I have drafted as Rmd files.

```{r}

wd = file.path( system.file(package = 'oetteR') )
folder_Rmd = 'POC_Rmd'
folder_html = 'POC_Rmd'

f_html_table_html_and_rmd_link( wd, folder_Rmd, folder_html )

```

# POC R

These are some proof of concepts that I have drafted out as plain R files

They can be found [here]( `r file.path( wd, 'POC_R' )` )

```{r}

R_files = tibble( file = dir( file.path( wd, 'POC_R' ), pattern = '\\.R', full.names = FALSE ) ) %>%
  mutate( path = dir( file.path( wd, 'POC_R'  ), pattern = '\\.R', full.names = TRUE )
          , link =  map_chr( path, function(x) f_html_filename_2_link( file_path = x ) )
          , file_prefix = str_split(file, '\\.')
          , file_prefix = map_chr( file_prefix, function(x) x[[1]] )
  ) %>%
  select(link)


f_datatable_minimal(R_files)

```

# POC Rmd Shiny

These are some proof of concepts for shiny functionality which I drafted as Rmd files.

They can be found [here]( `r file.path( wd, 'POC_Rmd_shiny' )` )

```{r}

R_files = tibble( file = dir( file.path( wd, 'POC_Rmd_shiny' ), pattern = '\\.R', full.names = FALSE ) ) %>%
  mutate( path = dir( file.path( wd, 'POC_Rmd_shiny'  ), pattern = '\\.R', full.names = TRUE )
          , link =  map_chr( path, function(x) f_html_filename_2_link( file_path = x ) )
          , file_prefix = str_split(file, '\\.')
          , file_prefix = map_chr( file_prefix, function(x) x[[1]] )
  ) %>%
  select(link)


f_datatable_minimal(R_files)

```


<br></br>
<br></br>
<br></br>
<br></br>
<br></br>
<br></br>
