---
title: "Visualising Regression Models"
author: "Bjoern Oettinghaus"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
  , eval = T
  , comment = "#>"
  , out.width = '100%'
  , message = F
  , warning = F
  , echo = T
)
```

```{r}
suppressPackageStartupMessages( require(oetteR) )
suppressPackageStartupMessages( require(tidyverse) )
```

# Introduction

Alluvial Plots can be a powerfull tool to visualise categorical data. It will group observations that have similar values across a set of dimensions and visualise them as flows. The individual flows can be emphasised through different colouring methods.

# Alluvial Plots

## Visualize Data in a tidy data format

For this dataformat the `f_plot_alluvial` function is suitable. Also see the help documentation and the examples of that function.

### Data

```{r}

data_ls = mtcars %>%
  f_clean_data()

data_tidy = data_ls$data
max_variables = 5
variables = c( data_ls$categoricals[1:3], data_ls$numericals[1:3] )

head(data_tidy, 10) %>%
  knitr::kable()

```

### Colouring

```{r}

f_plot_alluvial( data = data_tidy
                , variables = variables
                , max_variables = max_variables
                , fill_by = 'first_variable' )

f_plot_alluvial( data = data_tidy
                , variables = variables
                , max_variables = max_variables
                , fill_by = 'last_variable' )

f_plot_alluvial( data = data_tidy
                , variables = variables
                , max_variables = max_variables
                , fill_by = 'all_flows' )

f_plot_alluvial( data = data_tidy
                , variables = variables
                , max_variables = max_variables
                , fill_by = 'values' )

```

### Ordering

The order of the variables on the x axis is determined by the parameter variables. The order of any y values can be changed using the oder_levels argument. Simply pass the values you want to reorder as a character vector.

```{r}

f_plot_alluvial( data = data_tidy
                , variables = variables
                , max_variables = max_variables
                , fill_by = 'values'
                , order_levels = c('1', '0') )


```

## Visualize data in the 'gathered' format

### Data

Here we have more than one row for each observation and measurements that belong to the same group such as mean arrival delay is gathered in one column, which is indexed by the quarter column. In an alluvial Plot we might want to add another independent variable for coloring like in this case `carrier`.

```{r}

monthly_flights = nycflights13::flights %>%
  group_by(month, tailnum, origin, dest, carrier) %>%
  summarise() %>%
  group_by( tailnum, origin, dest, carrier) %>%
  count() %>%
  filter( n == 12 ) %>%
  select( - n ) %>%
  left_join( nycflights13::flights ) %>%
  .[complete.cases(.), ] %>%
  ungroup() %>%
  mutate( tailnum = pmap_chr(list(tailnum, origin, dest, carrier), paste )
          , qu = cut(month, 4)) %>%
  group_by(tailnum, carrier, origin, dest, qu ) %>%
  summarise( mean_arr_delay = mean(arr_delay) ) %>%
  ungroup() %>%
  mutate( mean_arr_delay = ifelse( mean_arr_delay < 10, 'on_time', 'late' ) )

levels(monthly_flights$qu) = c('Q1', 'Q2', 'Q3', 'Q4')

data_gath = monthly_flights

head(data_gath, 10) %>%
knitr::kable()

col_x = 'qu'
col_y = 'mean_arr_delay'
col_fill = 'carrier'
col_id = 'tailnum'


```


### Coloring

#### color by independent variable `carrier`

```{r}

f_plot_alluvial_1v1( data_gath, col_x = 'qu', col_y = 'mean_arr_delay', col_id = 'tailnum', col_fill )

```

##### Move colored independent variable to the left

```{r}

f_plot_alluvial_1v1( data_gath, col_x, col_y, col_id, col_fill, fill_right = F )

```


#### Other coloring options

```{r}

f_plot_alluvial_1v1( data_gath, col_x = 'qu', col_y = 'mean_arr_delay', col_id = 'tailnum', fill_by = 'last_variable' )
f_plot_alluvial_1v1( data_gath, col_x = 'qu', col_y = 'mean_arr_delay', col_id = 'tailnum', fill_by = 'first_variable' )
f_plot_alluvial_1v1( data_gath, col_x = 'qu', col_y = 'mean_arr_delay', col_id = 'tailnum', fill_by = 'all_flows' )
f_plot_alluvial_1v1( data_gath, col_x = 'qu', col_y = 'mean_arr_delay', col_id = 'tailnum', fill_by = 'value' )

```

### Ordering

#### Y levels of dependent variable

```{r}
f_plot_alluvial_1v1( data_gath, col_x, col_y, col_id, fill_by = 'first_variable'
                     , order_levels_y = c('on_time', 'late') )

```

#### Y levels of independent color variable `carrier`

```{r}

order_by_carrier_size = data_gath %>%
  group_by(carrier) %>%
  count() %>%
  arrange( desc(n) ) %>%
  .[['carrier']]

f_plot_alluvial_1v1( data_gath, col_x, col_y, col_id, col_fill
                     , order_levels_fill = order_by_carrier_size )


```


#### X levels

```{r}
f_plot_alluvial_1v1( data_gath, col_x, col_y, col_id, fill_by = 'first_variable'
                     , order_levels_x = c('Q4', 'Q3', 'Q2', 'Q1') )

```


## General

### Change the colors

Any color palette can be passed to both functions.

```{r}

f_plot_alluvial_1v1( data_gath, col_x, col_y, col_id, fill_by = 'last_variable'
                     , col_vector_flow = rev( RColorBrewer::brewer.pal(9, 'Purples') )
                     , col_vector_value = rev( RColorBrewer::brewer.pal(9, 'Oranges') ) )

f_plot_alluvial( data = data_tidy
                , variables = variables
                , max_variables = max_variables
                , fill_by = 'first_variable' 
                , col_vector_flow = rev( RColorBrewer::brewer.pal(9, 'Reds') )
                , col_vector_value = rev( RColorBrewer::brewer.pal(9, 'Greens') )
                )

```


### Connect Flows to observations in original data

```{r}

link = file.path( system.file(package = 'oetteR')
                  , 'Rmd vignettes'
                  , 'vignette_visualising_regression_models.html' )

```

The plot objects returned by both functions have an attribute called `data_key` which is an x-y table arranged like the alluvial plot one column containing the original ID. See [vignette for visualising regression models](`r link`) for an example on how this is effectively used.

#### tidy data

```{r}

p = f_plot_alluvial( data = data_tidy
                , variables = variables
                , max_variables = max_variables
                , fill_by = 'first_variable' )

p$data_key %>%
  head(10) %>%
  knitr::kable()

```

#### gathered data

```{r}

p = f_plot_alluvial_1v1( data_gath, col_x = 'qu', col_y = 'mean_arr_delay', col_id = 'tailnum', fill_by = 'last_variable' )


p$data_key %>%
  head(10) %>%
  knitr::kable()


```

### The Plot objects can be manipulated like any other ggplot object

#### Flip

```{r}

p = p +
  coord_flip()

p

```

